Intentional release of native species undermines ecological stability

### 2022-10-17 rmd/manuscript.Rmd
Update manuscript.Rmd

Massive releases of captive-bred native species ("intentional release") is a pervasive method to enhance wild populations of commercial and recreational species. However, such external inputs may disrupt the sensitive balance of species interactions that allows competing species to coexist, potentially compromising the long-term community stability. Here, we use theory and long-term data of stream fish communities to show that intentional release triggers community collapse with little demographic benefit to the enhanced species. Our theory predicted that intentional release intensifies interspecific competition, facilitating competitive exclusion of unenhanced species that otherwise stablely coexist. In parallel, the excessive input of captive-bred individuals suppressed the natural recruitment of the enhanced species via intensified within-species competition. Consequently, the ecological community with intentional release is predicted to show reduced community density with unstable temporal dynamics. Consistent with this prediction, stream fish communities showed greater temporal fluctuations and fewer species richness in rivers with the intensive release of hatchery salmon - a major fishery resource worldwide. Our findings alarm that the current overreliance on intentional release may accelerate global biodiversity loss, threatening the provisioning of ecosystem services.

@@

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) at sites with intensive release (**Figure 1B**). The effect was striking in its magnitude, almost doubling the CV at the highest release level. Our analysis strongly supported the positive relationship between the CV and the number of "effective" releases (**Figure 2**; see **Methods** for "effective" release), in which the probability of the regression coefficient being positive was `r op(df_beta %>% filter(response == "cv") %>% pull(prob), 2)` (**Table S6**). This pattern was associated with the reduced long-term average of the total community density and fewer taxonomic richness (**Figures 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures 1B and 2;** see **Tables S6-S8** for full statistics). In the meantime, the SDs had vague relationships with intentional release (**Figures 1B and 2**).

@@

Importantly, socioeconomic factors control the number of releases. The hatchery program aims to produce and release a constant number of hatchery masu salmon every year, although uncontrollable factors (e.g., diseases in a hatchery, budget allocation) generate considerable variation in release number among years. Therefore, the number of releases is determined without accounting for the current conditions of the recipient ecosystem. As such, released fish are probably "excessive" and may cause resource competition that would otherwise not exist.

It is noteworthy that the magnitude of the release program at the protected watersheds (annual average maximum = `r op(max(df_stock_mu$mean_stock), 2)` million fish) is comparable or even smaller when compared with those for other commercial and recreational species [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]. For example, 0.5 -- 1.2 million juveniles of Black sea bream are released annually in the Kagoshima bay, Japan, where substantial declines of genetic diversity and wild individuals have been reported [@kitadaRigorousMonitoringLargescale2019]. Similarly, three to four millions of Red-legged partridge are released for sports hunting in rural Spain [@villanuaSanitaryRisksRedlegged2008]. Therefore, our results should not be viewed as an extreme scenario. Importantly, a meta-analysis focusing on fishery resources revealed that only a few of these massive release programs led to increased abundance of the enhanced species [@arakiHatcheryStockingHelp2010]. Hence, the reduced abundance of enhanced species may be general, although none of those studies quantified the community-level impact of massive releases. Future studies addressing this knowledge gap is therefore highly valuable.

The near-intact environment of the protected watersheds with strict regulations of human activities (**Figure S10**) may have helped uncover the qualitative agreement between theory and empirical patterns. For example, any form of exploitation, including angling, is strictly prohibited for any fish species in the protected watersheds. In addition, an exclusive permission is required to physically alter in-stream and riparian habitats (e.g., construction of river-crossing structures). Therefore, unmeasured human influences should have minimal influences on the observed relationships. Further, we have statistically controlled natural variation in environmental factors (see **Methods**). Nevertheless, our results must be viewed with caution: we cannot exclude the possibility of spurious correlations due to the inherent nature of correlative analysis and field research. Since an experimental approach is nearly impossible at this spatial scale, the application of novel causal inferences (e.g., convergent cross mapping) [@sugiharaDetectingCausalityComplex2012; @ushioFluctuatingInteractionNetwork2018] may be a potential alternative to confirm the causal relationship between intentional release and community dynamics if sufficient time-series data are available.

Despite the significant attention to the fate of captive-bred individuals [@arakiHatcheryStockingHelp2010; @snyderLimitationsCaptiveBreeding1996], current schemes rarely consider the self-regulation process of biodiversity. Our results suggest that the ignorance of this critical process may erode the long-term persistence of the recipient community, likely impacting the stable delivery of ecosystem services [@cardinaleBiodiversityLossIts2012]. While more sophisticated analysis is required to provide detailed guidance on how to improve release programs, it is clear from our analysis that habitat conservation should be prioritized with a broader scope of ecosystem management for the sustainability of natural resources. Such efforts are particularly important in the Anthropocene because habitat degradation may exacerbate the undesired influence of intentional release by compromising the environmental capacity. Protected areas and environmental restoration are promising tools to conserve biodiversity, and a smart spatial design is a key to achieving successful conservation. For example, coordinated placement of conservation sites considering spatial biodiversity patterns is crucial in improving the ecological outcomes [@teruiEmergentDualScaling2021; @koningNetworkGrassrootsReserves2020; @fontouraProtectingConnectivityPromotes2022]. Governance may also play a central role in enforcing environmental legislation, potentially determining the effectiveness of conservation investment [@amanoSuccessfulConservationGlobal2018]. These considerable potentials indicate that viable management options exist before blindly accepting intentional release. Without a comprehensive framework that appreciates the ecological integrity of natural communities, the intentional release will never be effective but impairs biodiversity.

>>>>

Massive releases of captive-bred native species ("intentional release") is a pervasive method to enhance wild populations of commercial and recreational species. However, such external inputs may disrupt the sensitive balance of species interactions that allows competing species to coexist, potentially compromising the long-term community stability. Here, we use theory and long-term data of stream fish communities to show that intentional release destabilizes community dynamics with a limited demographic benefit to the enhanced species. Our theory predicted that intentional release intensifies interspecific competition, facilitating competitive exclusion of unenhanced species that otherwise stablely coexist. In parallel, the excessive input of captive-bred individuals suppressed the natural recruitment of the enhanced species via intensified within-species competition. Consequently, the ecological community with intentional release is predicted to show reduced community density with unstable temporal dynamics. Consistent with this prediction, stream fish communities showed greater temporal fluctuations and fewer species richness in rivers with the intensive release of hatchery salmon - a major fishery resource worldwide. Our findings alarm that the current overreliance on intentional release may accelerate global biodiversity loss, threatening the provisioning of ecosystem services.

@@

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) at sites with intensive release (**Figure 1B**). The effect was striking in its magnitude, almost doubling the CV at the highest release level. Our analysis strongly supported the positive relationship between the CV and the number of "effective" releases (**Figure 2**; see **Methods** for the Bayesian estimation of effective release), in which the probability of the regression coefficient being positive was `r op(df_beta %>% filter(response == "cv") %>% pull(prob), 2)` (**Table S6**). This pattern was associated with the reduced long-term average of the total community density and fewer taxonomic richness (**Figures 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures 1B and 2;** see **Tables S6-S8** for full statistics). In the meantime, the SDs had vague relationships with intentional release (**Figures 1B and 2**).

@@

Importantly, socioeconomic factors control the number of releases. The hatchery program aims to produce and release a constant number of hatchery masu salmon every year, although uncontrollable factors (e.g., diseases, budget allocation) generate considerable variation in the release number among years. Therefore, the number of releases is determined without accounting for the current conditions of the recipient ecosystem. As such, released fish are probably "excessive" and may cause resource competition that would otherwise not exist.

It is noteworthy that the magnitude of the release program at the protected watersheds (annual average maximum = `r op(max(df_stock_mu$mean_stock), 2)` million fish) is comparable or even smaller when compared with those for other commercial and recreational species [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]. For example, 0.5 -- 1.2 million juveniles of Black sea bream are released annually in the Kagoshima bay, Japan, where substantial declines of genetic diversity and wild individuals have been reported [@kitadaRigorousMonitoringLargescale2019]. Similarly, three to four millions of Red-legged partridge are released for sports hunting in rural Spain [@villanuaSanitaryRisksRedlegged2008]. Therefore, our results should not be viewed as an extreme scenario. We also must mention that a meta-analysis focusing on fishery resources revealed that only a few of massive release programs led to increased abundance of the enhanced species [@arakiHatcheryStockingHelp2010]. The reduced abundance of enhanced species may be general, although none of those studies statistically quantified the community-level impact of massive releases. Future studies addressing this knowledge gap is therefore highly valuable.

The near-intact environment of the protected watersheds with strict regulations of human activities (**Figure S10**) may have helped uncover the qualitative agreement between theory and empirical patterns. For example, any form of exploitation, including angling, is prohibited for any species in the protected watersheds. In addition, an exclusive permission is required to physically alter in-stream and riparian habitats (e.g., construction of river-crossing structures). Therefore, unmeasured human influences should have minimal influences on the observed relationships. Further, we have statistically controlled natural variation in environmental factors (see **Methods**). Nevertheless, our results must be viewed with caution: we cannot exclude the possibility of spurious correlations due to the inherent nature of correlative analysis and field research. Since an experimental approach is nearly impossible at this spatial scale, the application of novel causal inferences (e.g., convergent cross mapping) [@sugiharaDetectingCausalityComplex2012; @ushioFluctuatingInteractionNetwork2018] may be a potential alternative to confirm the causal relationship between intentional release and community dynamics if sufficient time-series data are available (a minimum of \> 30 temporal replicates).

Despite the significant attention to the fate of captive-bred individuals [@arakiHatcheryStockingHelp2010; @snyderLimitationsCaptiveBreeding1996], current schemes rarely consider the self-regulation process of biodiversity. Our results suggest that the ignorance of this critical process may erode the long-term persistence of the recipient community, likely impacting the stable delivery of ecosystem services [@cardinaleBiodiversityLossIts2012]. While socioeconomic analysis is required to provide detailed guidance on how to improve release programs, it is clear for the natural resources sustainability that habitat conservation should be prioritized with a broader scope of ecosystem management. Such efforts are particularly important in the Anthropocene because habitat degradation may facilitate the undesired influence of intentional release by compromising the environmental capacity. Protected areas and environmental restoration are promising tools to conserve biodiversity, and a smart spatial design is a key to achieving successful conservation. For example, coordinated placement of conservation sites considering spatial biodiversity patterns is crucial in improving the ecological outcomes [@akasakaSmartProtectedArea2017; @teruiEmergentDualScaling2021; @koningNetworkGrassrootsReserves2020; @fontouraProtectingConnectivityPromotes2022]. Governance may also play a central role in enforcing environmental legislation, potentially determining the effectiveness of conservation investment [@amanoSuccessfulConservationGlobal2018]. These considerable potentials indicate that viable management options exist before blindly accepting intentional release. Without a comprehensive framework that appreciates the ecological integrity of natural communities, the intentional release will never be effective but impairs biodiversity.
